---
title: "RainfallData_Kaggle"
author: "Manikandan R"
date: "3 October 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Context
#This data set contains monthly rainfall detail of 36 meteorological sub-divisions of India.

#Content

#Time Period: 1901 - 2015
#Granularity: Monthly
#Location: 36 meteorological sub-divisions in India
#Rainfall unit: mm

#Frame the Problem: Identify the key question are you trying to answer
# Rainfall data downloaded from the following link  https://www.kaggle.com/rajanand/rainfall-in-india 
# Which year registered highest rainfall ?
# Which Month registered highest rainfall ?
# Which subdivision registered highest rainfall for a given year ?
# Predict Rainfall for 2016 ad more months ?




Acquire the Data: Get the dataset to answer the question.
```{r}
getwd()
setwd("C:\\Users\\Iyshu\\Desktop\\GreatLakes\\Public data")
Rainfall <- read.csv(file = "rainfall in india 1901-2015.csv", header =  TRUE, sep =",")
```


```{r}
bak <- Rainfall
str(Rainfall)
colnames(Rainfall)

```

Refine the Data: Do the basic refinement to clean up the dataset. 
Seems no much refinement is needed for this dataset. :-)

```{r}
unique(Rainfall)
head(Rainfall)

Rainfall$SUBDIVISION <- as.character(Rainfall$SUBDIVISION)
str(Rainfall$SUBDIVISION)

```


```{r}
library(ggplot2)
library(tidyr)
library(rvest)
library(stringr)
library(plyr)
library(plotly)
library(tidyverse)

```
Transform the Data: Do the transformation needed for the dataset.
There is no much transfermation needed for this data :-)


```{r}
ChangeName <- c("ANDAMAN & NICOBAR ISLANDS"  = "A&N",         
      "ARUNACHAL PRADESH"            = "ACP",       
      "ASSAM & MEGHALAYA"           = "ASM",       
      "NAGA MANI MIZO TRIPURA"       = "TRP",      
      "SUB HIMALAYAN WEST BENGAL & SIKKIM" = "SIK",
      "GANGETIC WEST BENGAL"  = "GWB",              
      "ORISSA"                  = "ORI",           
      "JHARKHAND"               = "JRD",           
      "BIHAR"                   = "BIH",           
      "EAST UTTAR PRADESH"     = "EUP",            
      "WEST UTTAR PRADESH"     = "WUP",            
      "UTTARAKHAND"            = "UTD",            
      "HARYANA DELHI & CHANDIGARH" = "CHD",
      "PUNJAB"                       = "PNB",      
      "HIMACHAL PRADESH" = "HIP",
      "JAMMU & KASHMIR" =   "J&K",
      "WEST RAJASTHAN"  = "WRJ"  ,                    
      "EAST RAJASTHAN"  = "ERJ"  ,                 
      "WEST MADHYA PRADESH" = "WMP",               
      "EAST MADHYA PRADESH" = "EMP",               
      "GUJARAT REGION"= "GUJ",                     
      "SAURASHTRA & KUTCH" = "SAU",                
      "KONKAN & GOA"  = "GOA",                    
      "MADHYA MAHARASHTRA" = "MMH",                 
      "MATATHWADA" = "MAT",                         
      "VIDARBHA" = "VID" ,                          
      "CHHATTISGARH" = "CHT",                       
      "COASTAL ANDHRA PRADESH" = "CAP",             
      "TELANGANA" = "TEL",                           
      "RAYALSEEMA" = "RAY",                         
      "TAMIL NADU" = "TN",                         
      "COASTAL KARNATAKA" = "CKA",                  
      "NORTH INTERIOR KARNATAKA" = "NKA",           
      "SOUTH INTERIOR KARNATAKA" = "SKA",           
      "KERALA" = "KER",                             
      "LAKSHADWEEP" = "LAK") 


Rainfall$SUBDIVISION <- revalue(Rainfall$SUBDIVISION, ChangeName)
summary(is.na.data.frame(Rainfall))
unique(Rainfall$SUBDIVISION)
```

Explore the Data: Create the 3 - 4 individual visualisation that explore the dataset.


Findout NAs and Transform the data to appropriately

```{r}
Rainfall[!complete.cases(Rainfall),]
Rainfall$YEAR <- as.numeric(Rainfall$YEAR)
Rainfall$SUBDIVISION <- as.factor(Rainfall$SUBDIVISION)
```

Treat missing values in all the column 


```{r}
#install.packages("DMwR")
str(Rainfall)
library(DMwR)
knnOutput <- knnImputation(Rainfall[, !names(Rainfall) %in% "medv"])  # perform knn imputation.
anyNA(knnOutput)
```


```{r}
knnOutput[!complete.cases(knnOutput),]
head(knnOutput)
unique(Rainfall$SUBDIVISION)
```

```{r}
summary(is.na.data.frame(knnOutput))
str(knnOutput)

```
Explore the Data: Create the 3 - 4 individual visualisation that explore the dataset.

```{r}

ggplot(data=knnOutput, aes(x=knnOutput$YEAR, y = knnOutput$ANNUAL, group=knnOutput$ANNUAL, fill = knnOutput$YEAR                           )) +
    geom_line() +
    geom_point()

```

```{r}
ggplot(data=knnOutput, aes(x=YEAR
                           , y=ANNUAL, fill=SUBDIVISION)) +
    geom_bar(stat="identity",position=position_dodge(),color="Red") 
#facet_wrap( ~ ANNUAL, ncol=3) 
```



```{r}


ggplot(data=knnOutput) + geom_bar(aes(x=reorder(SUBDIVISION, -ANNUAL), y=ANNUAL),stat="identity",width=0.5) +
  ylab("Annual rainfall in mm") +
  ggtitle("Subdivision received rainfall") +
  theme(axis.text.x = element_text(angle=60, hjust=1,size=7), aspect.ratio = 0.9,axis.ticks = element_blank()) 
  #coord_flip()

  
```


```{r}

newData <- knnOutput
head(newData)
knnOutput$ANNUAL<- as.numeric(knnOutput$ANNUAL)
str(knnOutput)
knnOutput <- knnOutput %>% 
             #group_by(knnOutput$SUBDIVISION) %>%
		         select(knnOutput$SUBDIVISION,knnOutput$ANNUAL)

head(10)
```



```{r}
max(knnOutput$YEAR)
min(knnOutput$YEAR)
```

```{r}
annual<-newData[,1]
head(annual)
a=ts(annual,start=1800,end=2000,frequency = 12)
a
```



```{r}
plot.ts(a)
```
```{r}
plot(knnOutput$ANNUAL ~ knnOutput$YEAR, type="h", xlab="Year", ylab="Annual rainfail in mm", col="blue")

```


```{r}

head(knnOutput)
knnOutput$SUBDIVISION <- as.factor(knnOutput$SUBDIVISION)
knnOutput %>% group_by(YEAR,SUBDIVISION) %>%
           ts(start=c(1901,1), freq=1) -> rain_ts 
          rain_ts %>% window(c(1901,1),c(2014,1))
          
```
```{r}

ggplot(data=knnOutput) +
   geom_bar(aes(x=SUBDIVISION, y=ANNUAL), stat="identity",width = 0.5)+
  ylab("Annual rainfall in mm") +
  ggtitle("Location ") +
  theme(axis.text.x = element_text(angle=60, hjust=1, size=7), aspect.ratio=0.9, axis.ticks=element_blank())
```



Lets see the rainfall data with the interactive graph.


```{r}
#install.packages("dygraphs")
library(dygraphs) # A dynamic graph library
rain_ts %>% dygraph(width=600,height=300) %>% dyRangeSelector


```


```{r}
rain_ts %>% dygraph(width=600,height=300) %>% dyRangeSelector %>% dyRoller(rollPeriod = 600)

```


```{r}
hrf <- newData
```

```{r}

quarter <- hrf[,c(-3:-14)]
quarter$SUBDIVISION <- as.character(quarter$SUBDIVISION)
str(quarter)
plot(lm(formula = YEAR ~ ANNUAL, data=quarter))




```



```{r}
barplot(quarter$ANNUAL, quarter$YEAR , col="blue", lwd=2,
        xlab = "Year of rainfall", ylab="Annual rain in mm", border=par("fg"))
```
```{r}
ha = HoltWinters(a,beta=F,gamma=F)
ha
ha$fitted
ha$SSE
```

Prediction for next 100 months with the available data


```{r}
library(forecast)
fha =forecast:::forecast.HoltWinters(ha,h=100)
fha
forecast:::plot.forecast(fha)
```

The above analysis answers some of the questions defined initially.


